---
title: "Dow Jones Case Study"
author: "Charles Reyes"
date: "10/31/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/Krischelle/Documents/Krischelles most important stuff/Publish/Books and Learning/UTSA/Fall 2021 Classes/Data Analytics Applications/Case Studies/Dow Jones Case Study/")
library(ggplot2); library(reshape2); library(dplyr); library(wesanderson); library(tidyverse); library(e1071); library(tree); library(quantmod)
```

Reading data

```{r}
dow_jones_1 = read.csv("dow_jones_index.data", header = TRUE)

dow_jones_2 = dow_jones_1

DJIDta = read.csv("DJI.csv", header = TRUE, sep = ",")
```

Viewing structure of data.

```{r}
str(dow_jones_1)
``` 

Cleaning data:

- Converting all currency variables from text to numeric
- Converting `date` variable from text to date
- Converting `stock` variable from character to factor

```{r}
dow_jones_1$open = as.numeric(gsub("\\$", "", dow_jones_1$open))
dow_jones_1$high = as.numeric(gsub("\\$", "", dow_jones_1$high))
dow_jones_1$low = as.numeric(gsub("\\$", "", dow_jones_1$low))
dow_jones_1$close = as.numeric(gsub("\\$", "", dow_jones_1$close))
dow_jones_1$next_weeks_open = as.numeric(gsub("\\$", "", dow_jones_1$next_weeks_open))
dow_jones_1$next_weeks_close = as.numeric(gsub("\\$", "", dow_jones_1$next_weeks_close))
dow_jones_1$date = as.Date(dow_jones_1$date, "%m/%d/%Y")
dow_jones_1$stock = as.factor(dow_jones_1$stock)
```

```{r}
str(dow_jones_1)
```

Viewing summary of data

```{r}
summary(dow_jones_1)
```

Sorting data by `stock` and `date`

```{r}
dow_jones_1 = dow_jones_1[order(dow_jones_1$stock, dow_jones_1$date), ]
```

Removing rows with null values to check correlation

```{r}
dow_jones_new = dow_jones_1[dow_jones_1$date != as.Date("2011-01-07"), ]
```

Checking correlation of data

```{r}
cor(dow_jones_new[ , 4:16])
```

Creating heat map to visualize correlation

```{r}
dow_jones_cor = melt(round(cor(dow_jones_new[ , 4:16]), 2))
ggplot(data = dow_jones_cor, aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile() +
  scale_fill_gradientn(colours = wes_palette("Zissou1", 100, type = "continuous")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Removing highly correlated variables not needed for analysis

```{r}
dow_jones_new2 = dow_jones_new[ , -c(1, 4:6, 11:13)]
str(dow_jones_new2)
```

Lag plots for remaining highly correlated variable

```{r}
lag.plot(dow_jones_new2$close, pch = ".", set.lags = 1:4)
```

Lag 1 looks to be the best

Creating lag 1 variables in data set

```{r}
dow_jones_lag = dow_jones_new2 %>%                            
  group_by(stock) %>%
  dplyr::mutate(close_lag1 = lag(close, n = 1, default = NA))
```

Removing rows with null values to check correlation

```{r}
dow_jones_lag_new = dow_jones_lag[dow_jones_lag$date != as.Date("2011-01-07"), ]
dow_jones_lag_new = dow_jones_lag_new[dow_jones_lag_new$date != as.Date("2011-01-14"), ]
```

Creating heat map to visualize correlation again after removing variables

```{r}
dow_jones_lag_cor = melt(round(cor(dow_jones_lag_new[ , 3:10]), 2))
ggplot(data = dow_jones_lag_cor, aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile() +
  scale_fill_gradientn(colours = wes_palette("Zissou1", 100, type = "continuous")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Creating final data set to use for analysis

```{r}
dow_jones_final = dow_jones_1[ , -c(4:6, 11:13)]

dow_jones_final = dow_jones_final %>%                            
  group_by(stock) %>%
  dplyr::mutate(close_lag1 = lag(close, n = 1, default = NA))
```

```{r}
str(dow_jones_final)
```

Splitting data into train and test data sets

```{r}
train = dow_jones_final[dow_jones_final$quarter == 1, ]
# Removing quarter column
train = train[ , -1]

# not sure if we should remove 1st row of data since it has null values; leaving in for now
# train = train[train$date != as.Date("2011-01-07"), ]

test = dow_jones_final[dow_jones_final$quarter == 2, ]
# Removing quarter column
test = test[ , -1]

train1 = split(train, train$stock)

test1 = split(test, test$stock)
```

# Linear Regression model

```{r}
glmfxn = function (trainstock, teststock, formula) {
    set.seed(123)
    glmfit = glm(formula, data = trainstock)
    linear.predict = predict.glm(glmfit, newdata = teststock)
    linear.predict = ifelse(linear.predict >= mean(linear.predict), "1", "0")
    linear.actual = ifelse(teststock$percent_change_next_weeks_price >= mean(teststock$percent_change_next_weeks_price), "1", "0")

  confusion = caret::confusionMatrix(as.factor(linear.actual), as.factor(linear.predict))

    print(confusion$overall)
    print(summary(glmfit))

  }

formula = percent_change_next_weeks_price ~ close_lag1 + percent_return_next_dividend + volume 

for (l in names(train1))

{
  x = train1[[l]]
  y = test1[[l]]
 glmfxn(x, y, formula)
  namefilmodel = paste0("Linear_Accuracy ", l)
  print(namefilmodel)
  
  }
```

# Decision Tree

```{r}
treefxn = function(trainstock, teststock, formula){
  set.seed(123)
 
  treefit = tree(formula, data = trainstock)
 
  tree.predict = predict(treefit, newdata = teststock)
 
  tree.predict = ifelse(tree.predict >= mean(tree.predict), 1, 0)
 
  tree.obs = ifelse(teststock$percent_change_next_weeks_price >= mean(teststock$percent_change_next_weeks_price ), 1, 0)
  table = table(tree.obs, tree.predict); print(summary(treefit))
 
  accuracy = sum(diag(table))/12
 
  print(accuracy)
}

formula = percent_change_next_weeks_price ~ close_lag1 + percent_return_next_dividend + volume

for (i in names(train1)){
  x = train1[[i]]
  y = test1[[i]]
  treefxn(x, y, formula)
}
```

# SVR Radial Model

```{r}
svmfxn = function(trainstock, teststock, formula){
  set.seed(123)
 
  svmfit = svm(formula, data = trainstock, kernerl = "radial", cost = 0.1, gamma = 0.01)
 
  svm.predict = predict(svmfit, newdata = teststock, type = "class")
 
  svm.predict = ifelse(svm.predict >= mean(svm.predict), 1, 0)
 
  svm.obs = ifelse(teststock$percent_change_next_weeks_price >= mean(teststock$percent_change_next_weeks_price ),1,0)
  table = table(svm.obs, svm.predict)
 
  accuracy = sum(diag(table))/12
  print(summary(svmfit))
  print(accuracy)
}

formula = percent_change_next_weeks_price ~ close_lag1 + percent_return_next_dividend + volume

for (i in names(train1)){
  x= train1[[i]]
  y= test1[[i]]
  svmfxn(x, y, formula)
}
```

# CAPM and stock risk and reward calculations

Numeric conversion
```{r}
dow_jones_2$open=as.numeric(gsub("\\$","",dow_jones_2$open))
dow_jones_2$high=as.numeric(gsub("\\$","",dow_jones_2$high))
dow_jones_2$low=as.numeric(gsub("\\$","",dow_jones_2$low))
dow_jones_2$close=as.numeric(gsub("\\$","",dow_jones_2$close))
dow_jones_2$next_weeks_open=as.numeric(gsub("\\$","",dow_jones_2$next_weeks_open))
dow_jones_2$next_weeks_close=as.numeric(gsub("\\$","",dow_jones_2$next_weeks_close))

DowJData3 <- split(dow_jones_2, dow_jones_2$stock)

ReturnDJI = na.omit(DJIDta[,5]) 

ReturnAA = na.omit(Delt(DowJData3$AA[,7])); ReturnAA
```

```{r}
ReturnAXP = na.omit(Delt(DowJData3$AXP[,7]))
ReturnBA = na.omit(Delt(DowJData3$BA[,7]))
ReturnBAC = na.omit(Delt(DowJData3$BAC[,7]))
ReturnCAT = na.omit(Delt(DowJData3$CAT[,7]))
ReturnCSCO = na.omit(Delt(DowJData3$CSCO[,7]))
ReturnCVX = na.omit(Delt(DowJData3$CVX[,7]))
ReturnDD = na.omit(Delt(DowJData3$DD[,7]))
ReturnDIS = na.omit(Delt(DowJData3$DIS[,7]))
ReturnGE = na.omit(Delt(DowJData3$GE[,7]))
ReturnHD = na.omit(Delt(DowJData3$HD[,7]))
ReturnHPQ = na.omit(Delt(DowJData3$HPQ[,7]))
ReturnIBM = na.omit(Delt(DowJData3$IBM[,7]))
ReturnINTC = na.omit(Delt(DowJData3$INTC[,7]))
ReturnJNJ = na.omit(Delt(DowJData3$JNJ[,7]))
ReturnJPM = na.omit(Delt(DowJData3$JPM[,7]))
ReturnKO = na.omit(Delt(DowJData3$KO[,7]))
ReturnKRFT = na.omit(Delt(DowJData3$KRFT[,7]))
ReturnMCD = na.omit(Delt(DowJData3$MCD[,7]))
ReturnMMM = na.omit(Delt(DowJData3$MMM[,7]))
ReturnMRK = na.omit(Delt(DowJData3$MRK[,7]))
ReturnMSFT = na.omit(Delt(DowJData3$MSFT[,7]))
ReturnPFE = na.omit(Delt(DowJData3$PFE[,7]))
ReturnPG = na.omit(Delt(DowJData3$PG[,7]))
ReturnT = na.omit(Delt(DowJData3$T[,7]))
ReturnTRV = na.omit(Delt(DowJData3$TRV[,7]))
ReturnUTX = na.omit(Delt(DowJData3$UTX[,7]))
ReturnWMT= na.omit(Delt(DowJData3$WMT[,7]))
ReturnVZ = na.omit(Delt(DowJData3$VZ[,7]))
ReturnXOM = na.omit(Delt(DowJData3$XOM[,7]))
```

```{r}
MyData = cbind(ReturnDJI,
               ReturnAA,
               ReturnAXP,
               ReturnBA, 
               ReturnBAC,
               ReturnCAT,
               ReturnCSCO,
               ReturnCVX,
               ReturnDD,
               ReturnDIS,
               ReturnGE,
               ReturnHD,
               ReturnHPQ,
               ReturnIBM,
               ReturnINTC,
               ReturnJNJ,
               ReturnJPM,
               ReturnKO,
               ReturnKRFT,
               ReturnMCD,
               ReturnMMM,
               ReturnMRK,
               ReturnMSFT,
               ReturnPFE,
               ReturnPG,
               ReturnT,
               ReturnTRV,
               ReturnUTX,
               ReturnWMT,
               ReturnVZ,
               ReturnXOM)
               
colnames(MyData) = c("DJI", "AA", "AXP", "BA", "BAC", "CAT", "CSCO", "CVX", "DD", "DIS", "GE", "HD", "HPQ", "IBM", "INTC", "JNJ", "JPM", "KO", "KRFT", "MCD", "MMM", "MRK", "MSFT", "PFE", "PG", "T", "TRV", "UTX", "WMT", "VZ", "XOM")
head(MyData)
```

```{r}
str(MyData)
```

Compute mean and standard deviation

```{r}
DataMean = apply(MyData, 2, mean)
DataSD = apply(MyData, 2, sd)

cbind(DataMean, DataSD)
```

```{r}
#1
lm.AA <- lm(AA~ DJI, data = as.data.frame(MyData))
#2
lm.AXP <- lm(AXP~ DJI, data = as.data.frame(MyData))
#3
lm.BA <- lm(BA~ DJI, data = as.data.frame(MyData))
#4
lm.BAC <- lm(BAC~ DJI, data = as.data.frame(MyData))
#5
lm.CAT <- lm(CAT~ DJI, data = as.data.frame(MyData))
#6
lm.CSCO <- lm(CSCO~ DJI, data = as.data.frame(MyData))
#7
lm.CVX <- lm(CVX~ DJI, data = as.data.frame(MyData))
#8
lm.DD <- lm(DD~ DJI, data = as.data.frame(MyData))
#9
lm.DIS <- lm(DIS~ DJI, data = as.data.frame(MyData))
#10
lm.GE <- lm(GE~ DJI, data = as.data.frame(MyData))
#11
lm.HD <- lm(HD~ DJI, data = as.data.frame(MyData))
#12
lm.HPQ <- lm(HPQ~ DJI, data = as.data.frame(MyData))
#13
lm.IBM <- lm(IBM~ DJI, data = as.data.frame(MyData))
#14
lm.INTC <- lm(INTC~ DJI, data = as.data.frame(MyData))
#15
lm.JNJ <- lm(JNJ~ DJI, data = as.data.frame(MyData))
#16
lm.JPM <- lm(JPM~ DJI, data = as.data.frame(MyData))
#17
lm.KO <- lm(KO~ DJI, data = as.data.frame(MyData))
#18
lm.KRFT <- lm(KRFT~ DJI, data = as.data.frame(MyData))
#19
lm.MCD <- lm(MCD~ DJI, data = as.data.frame(MyData))
#20
lm.MMM <- lm(MMM~ DJI, data = as.data.frame(MyData))
#21
lm.MRK <- lm(MRK~ DJI, data = as.data.frame(MyData))
#22
lm.MSFT <- lm(MSFT~ DJI, data = as.data.frame(MyData))
#23
lm.PFE <- lm(PFE~ DJI, data = as.data.frame(MyData))
#24
lm.PG <- lm(PG~ DJI, data = as.data.frame(MyData))
#25
lm.T <- lm(T~ DJI, data = as.data.frame(MyData))
#26
lm.TRV <- lm(TRV~ DJI, data = as.data.frame(MyData))
#27
lm.UTX <- lm(UTX~ DJI, data = as.data.frame(MyData))
#28
lm.WMT <- lm(WMT~ DJI, data = as.data.frame(MyData))
#29
lm.VZ <- lm(VZ~ DJI, data = as.data.frame(MyData))
#30
lm.XOM <- lm(XOM~ DJI, data = as.data.frame(MyData))
```

```{r}
BetaDowJ = summary(lm.AA)$coefficient[2,1]
paste("Beta of AA:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.AXP)$coefficient[2,1]
paste("Beta of AXP:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.BA)$coefficient[2,1]
paste("Beta of BA:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.BAC)$coefficient[2,1]
paste("Beta of BAC:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.CAT)$coefficient[2,1]
paste("Beta of CAT:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.CSCO)$coefficient[2,1]
paste("Beta of CSCO:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.CVX)$coefficient[2,1]
paste("Beta of CVX:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.DD)$coefficient[2,1]
paste("Beta of DD:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.DIS)$coefficient[2,1]
paste("Beta of DIS:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.GE)$coefficient[2,1]
paste("Beta of GE:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.HD)$coefficient[2,1]
paste("Beta of HD:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.HPQ)$coefficient[2,1]
paste("Beta of HPQ:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.IBM)$coefficient[2,1]
paste("Beta of IBM:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.INTC)$coefficient[2,1]
paste("Beta of INTC:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.JNJ)$coefficient[2,1]
paste("Beta of JNJ:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.JPM)$coefficient[2,1]
paste("Beta of JPM:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.KO)$coefficient[2,1]
paste("Beta of KO:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.KRFT)$coefficient[2,1]
paste("Beta of KRFT:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.MCD)$coefficient[2,1]
paste("Beta of MCD:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.MMM)$coefficient[2,1]
paste("Beta of MMM:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.MRK)$coefficient[2,1]
paste("Beta of MRK:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.MSFT)$coefficient[2,1]
paste("Beta of MSFT:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.PFE)$coefficient[2,1]
paste("Beta of PFE:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.PG)$coefficient[2,1]
paste("Beta of PG:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.T)$coefficient[2,1]
paste("Beta of T:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.TRV)$coefficient[2,1]
paste("Beta of TRV:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.UTX)$coefficient[2,1]
paste("Beta of UTX:" , BetaDowJ)
```

```{r}

BetaDowJ <- summary(lm.WMT)$coefficient[2,1]
paste("Beta of WMT:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.VZ)$coefficient[2,1]
paste("Beta of VZ:" , BetaDowJ)
```

```{r}
BetaDowJ <- summary(lm.XOM)$coefficient[2,1]
paste("Beta of XOM:" , BetaDowJ)
```



